---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ploom

[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/ploom)](https://cran.r-project.org/package=ploom)
[![Build Status](https://api.travis-ci.com/blakeboswell/ploom.svg?branch=develop)](https://api.travis-ci.com/blakeboswell/ploom)
[![Coverage status](https://codecov.io/gh/blakeboswell/ploom/branch/develop/graph/badge.svg)](https://codecov.io/github/blakeboswell/ploom?branch=develop)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

<!-- [AppVeyor Build Status]() -->

## Overview

`ploom` provides tools for __memory efficient__ fitting of Linear and Generalized Linear models. `ploom` fits models with a [bounded memory](#aknowledgements) aglorithm enabling:

- __Out-of-memory__ processing capable of fitting billions of observations
- __Bounded in-memory__ processing with runtimes comparable to `lm()` and `glm()`

`ploom` models are compatible with many [`stats`]() and [`tidy`]() statistical analysis functions. They are capable of generating robust standard errors, confidence and prediction intervals, residuals, and other artifacts for inferential analysis.

## Installation

```{r, eval = FALSE, message=FALSE}
# development version from GitHub:
# install.packages("devtools")
devtools::install_github("blakeboswell/ploom")
```

## Usage

Models are intialized with a `formula()`; fit to data with calls to `fit()`; and summarized with standard functions such as `tidy()`, `glance()`, and `summary()`.

```{r, eval = TRUE, message=FALSE}
library(ploom)

y <- oomlm(mpg ~ cyl + disp)
y <- fit(y, data = mtcars)

tidy(y)
```

#### Fitting in Chunks

Models can be be fit with repeated calls to `fit()` over chunks of data. Each call to `fit()` only needs to allocate memory for the provided chunk, thereby bounding the fitting process to chunk size.

The function `oomdata_tbl()` enables iteration over an in-memory `tibble()` or `data.frame()` object.

```{r, eval = TRUE, message=FALSE}
chunks <- oomdata_tbl(mtcars, chunk_size = 16)

while((!is.null(chunk <- chunks()))) {
  print(nrow(chunk))
}
```

When provided to `fit()`, all chunks will be iterated over and fit.

```{r, eval = TRUE, message=FALSE}
y <- fit(oomlm(mpg ~ cyl + disp), chunks)

# glance(y)
```

#### Working with Databases

Similarly, `oomdata_dbi()` facilitaties fitting to data in a database.  It is compatible with any database having a `DBI` backend.

```{r, eval = FALSE, message=FALSE}
#' connect to database
con     <- DBI::dbConnect(RPostgres::Postgres(), dbname = "mtcars")
result  <- DBI::dbSendQuery(con, "select mpg, cyl, disp from mtcars;")

#' fit model
X <- oomdata_dbi(result, chunk_size = 1)
y <- fit(oomlm(mpg ~ cyl + disp), X)
```

See the articles [NA]() and [NA]() for more on interfacing with databases.

#### Inference and Prediction

## Alternatives

- [`biglm`](https://cran.r-project.org/web/packages/biglm/index.html)
- [`speedglm`](https://cran.r-project.org/web/packages/speedglm/index.html)

## Acknowledgements

Thanks to:

- Fitting is performed via Alan Miller’s AS274 updating QR factorization algorithm
- [`biglm`](https://cran.r-project.org/web/packages/biglm/index.html)

