---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ploom

[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/ploom)](https://cran.r-project.org/package=ploom)
[![Build Status](https://api.travis-ci.com/blakeboswell/ploom.svg?branch=develop)](https://api.travis-ci.com/blakeboswell/ploom)
[![Coverage status](https://codecov.io/gh/blakeboswell/ploom/branch/develop/graph/badge.svg)](https://codecov.io/github/blakeboswell/ploom?branch=develop)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

<!-- [AppVeyor Build Status]() -->

# Overview

ploom provides tools for _fast_ and __memory efficient__ fitting of Linear and Generalized Linear models.

Features of ploom include:

- Out-of-memory processing capable of fitting __billions__ of observations.
- _Fast_ in-memory processing requiring less resources than `lm()` and `glm()`.
- Compatible with `tidy()`, `glance()`, `augment()` and many stats functions such as `predict()` and `residuals()`.
- Provides functions for iteratively processing data stored in-memory, in a database or on disk.

## Installation

```{r, eval = FALSE, message=FALSE}
# development version from GitHub:
# install.packages("devtools")
devtools::install_github("blakeboswell/ploom")
```

## Usage

Models are intialized with a `formula`; fit to data with calls to `fit()`; and summarized with standard functions such as `tidy()`, `glance()`, and `summary()`.

```{r, eval = TRUE, message=FALSE}
library(ploom)

y <- oomlm(mpg ~ wt + qsec + factor(am))
y <- fit(y, data = mtcars)

tidy(y)

```

### Bounded Memory

Models can be be fit with repeated calls to `fit()` over chunks of data. Each call to `fit()` only needs to allocate memory for the provided chunk, thereby bounding the required memory.

```{r, eval = TRUE, message=FALSE}
# create model
y <- oomlm(mpg ~ wt + qsec + factor(am))

# fit data in two chunks
y <- fit(y, mtcars[1:16, ])
y <- fit(y, mtcars[17:32, ])

coef(y)
```

### Iterative Fitting

The function `oomdata_tbl()` enables iteration over an in-memory `tibble`. When an `oomdata_tbl()` is provided as the data argument to `fit()`, all chunks are automatically iterated over.

```{r, eval = TRUE, message=FALSE}
chunks <- oomdata_tbl(mtcars, chunk_size = 16)
fit(oomlm(mpg ~ wt + qsec + factor(am)), chunks)
```

### Working with Databases

The function `oomdata_dbi()` enables iteratation over a [`DBI`]() result set. `fit()` will automatically fit the model over all chunks.

```{r, eval = TRUE, message=FALSE}
# connect to database
con    <- DBI::dbConnect(RPostgres::Postgres(), dbname="mtcars")
result <- DBI::dbSendQuery(con, "select mpg, wt, qsec, am from mtcars;")
chunks <- oomdata_dbi(result, chunk_size = 16)

# fit model to all chunks
y <- fit(oomlm(mpg ~ wt + qsec + factor(am)), chunks)

# inspect fit statistiscs
glance(y)
```

See the articles [NA]() and [NA]() for more on interfacing with databases.

### Prediction & Residuals

Prediction with ploom models is performed with the `predict()` function. Because ploom models do not store any data while fitting, we must also provide data. 

```{r, eval = TRUE, message=FALSE}
predict(y, new_data = mtcars, std_error = TRUE, interval = "prediction")
```

Similarly, residuals are accessible on demand with `residuals()`:

```{r, eval = TRUE, message=FALSE}
sum(residuals(y, data = mtcars)^2)
```

## Alternatives

- [`biglm`](https://cran.r-project.org/web/packages/biglm/index.html)
- [`speedglm`](https://cran.r-project.org/web/packages/speedglm/index.html)

## Acknowledgements

Thanks to:

- Fitting is performed via Alan Miller’s AS274 updating QR factorization algorithm
- [`biglm`](https://cran.r-project.org/web/packages/biglm/index.html)

