---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ploom

<!-- [CRAN_Status_Badge]() -->
<!-- [Build Status]() -->
<!-- [AppVeyor Build Status]() -->
<!-- [Coverage Status]() -->

## Overview

A collection of tools for out-of-memory linear model fitting and inference. Implements `lm` and `glm` analogs using Alan Miller's AS274 updating QR factorization algorithm. Collects and reports an array of pertinent fit statistics. Provides flexible and easy to use mechanisms to stream in data and stream out results during fitting.

> Currently in early development stage.

### tl;dr Features

> forthcoming

## Installation

```{r, eval = FALSE, message=FALSE}
# the early development version from GitHub:
# install.packages("devtools")
devtools::install_github("bboswell/ploom")
```

## Usage

The core of `ploom` consists of

- __`oomlm`__ and __`oomglm`__ for fitting linear and generalized linear models to data
- __`oomfeeds`__ that provide a flexible interface for streaming data to and from `oomlm` and `oomglm`

```{r, eval = TRUE, include=FALSE}
library(ploom)
```

The functions `oomlm` and `oomglm` work as you would expect when fitting in-memory data.

```{r, eval = TRUE}
w <- oomlm(mpg ~ cyl + disp, data = mtcars)
```

Similar to `biglm`, `ploom` can be updated with new data after being initially fit.

```{r, eval = TRUE}
# proxy for big data feed
chunks <- purrr::pmap(mtcars, list)

# initial fit
x  <- oomlm(mpg ~ cyl + disp, chunks[[1]])

# iteratively update model with more data
for(chunk in chunks[2:length(chunks)]) {
  x <- update_oomlm(x, data = chunk)
}
```

`ploom` models can also be initialized with a formula only, providing flexibility for initial updates.

```{r, eval = TRUE}
# no index loop
y <- oomlm(mpg ~ cyl + disp)
for(chunk in chunks) {
  update_oomlm(y, chunk)
}

# or avoiding loops altogether with `reduce`
z <- purrr::reduce(
  chunks, update_oomlm, .init = oomlm(mpg ~ cyl + disp)
  )
```


### `streams`

```{r, eval = FALSE}

```


## Alternatives

> forthcoming

## Acknowledgements

> forthcoming