---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ploom

<!-- [CRAN_Status_Badge]() -->
[![Build Status](https://api.travis-ci.com/blakeboswell/ploom.svg?branch=develop)](https://api.travis-ci.com/blakeboswell/ploom)
<!-- [AppVeyor Build Status]() -->
<!-- [Coverage Status]() -->

# Overview

A collection of tools for out-of-memory linear model fitting and inference. Implements `lm` and `glm` analogs using Alan Miller's AS274 updating QR factorization algorithm. Collects and reports an array of pertinent fit statistics. Provides flexible and easy to use mechanisms to stream in data and stream out results during fitting.

> Currently in early development stage.

## Features

> forthcoming

# Installation

```{r, eval = FALSE, message=FALSE}
# the early development version from GitHub:
# install.packages("devtools")
devtools::install_github("blakeboswell/ploom")
```

# Usage

See documentation.

# Alternatives

> forthcoming

# Acknowledgements

> forthcoming